---
title: "Homework 8"
author: "Mei Jiao"
date: "10/31/2020"
output: 
  pdf_document:
   dev: png

---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=2,echo=TRUE, warning=FALSE, message=FALSE)
```

# 5.3.3 Orstein–Uhlenbeck Process

Consider the Ornstein-Uhlenbeck process
\begin{align*}
  d r(t) = \alpha(b - r(t))dt + \sigma d W(t)
\end{align*}
where $\alpha > 0$, $\sigma > 0$, and $b$ are constants.
 
## 1. Show that for t > 0, $\Delta > 0$,

\begin{align*}
r(t+\Delta) = e^{-\alpha\Delta}r(t) + b(1-e^{-\alpha\Delta})+ \frac{\sigma}{\sqrt{2\alpha}}\sqrt{1-e^{-2\alpha\Delta}}Z
\end{align*}

## where Z~N(0,1)

Let \[V(t)=e^{\alpha t}(r(t)-b)\] we can get
\[\frac{\partial V}{\partial t} = \alpha e^{\alpha t}(r(t)-b)\]
\[\frac{\partial V}{\partial r} = e^{\alpha t}\]
\[\frac{\partial^{2} V}{\partial r^{2}}=0\]
then $dV$ can be expressed as
\[dV=\frac{\partial V}{\partial t}dt+\frac{\partial V}{\partial r}dr+\frac{1}{2}\frac{\partial^{2} V}{\partial r^{2}}dr^{2}\]
simplify it as
\[dV = \sigma e^{\alpha t}dW(t)\]
so\[V(T)-V(t)=e^{\alpha T}(r(T)-b)-e^{\alpha t}(r(t)-b)=\int_{t}^{T}\sigma e^{\alpha u}dW(u) \]
divided by $e^{\alpha T}$
\[r(T)-b -e^{\alpha (-T+t)}(r(t)-b)=\int_{t}^{T}\sigma e^{\alpha u-\alpha T}dW(u)\]
and we've known that
\[\int_{t}^{T}\sigma e^{\alpha u-\alpha T}dW(u)=\frac{\sigma}{\sqrt{2\alpha}}\sqrt{1-e^{-2\alpha (T-t)}} Z\]
so the solution is
\[r(T)=b(1-e^{e^{\alpha (-T+t)}})+e^{\alpha (-T+t)} r(t)+\frac{\sigma}{\sqrt{2\alpha}}\sqrt{1-e^{-2\alpha (T-t)}} Z\]

## 2. Use the transition distribution from the last part to implement a random walk construction for the process on time interval $[0,T]$. Your code should take $\alpha$, $\sigma$, $b$ the initial value $r(0)$, T, and the time step $\Delta$ of the random walk as input arguments. For $r(0)=1$, $T=500$ and $\Delta = 1/500$, plot a sample path for each combination of the following values,
  \begin{align*}
    \alpha \in \{0.1, 1, 5\},\ \sigma \in \{0.1, 0.2, 0.5\},\
    b\in\{-5, 5\}.
  \end{align*}

## Comment on how the behavior of $r(t)$ depends on $\alpha$ and $\sigma$.

First, we need to get the data, then to plot graphs to find out the behavior of r dependents on $\alpha$ and $\sigma$. The result shows that with the increasing of the $\alpha$, the path will be more quickly to become stable, and with the increase of the $sigma$, the wave will become lager. 

```{r}
r_t <- function(r,a,b,sigma,delta){
  exp(-a*delta)*r+b*(1-exp(-a*delta))+sigma/sqrt(2*a)*sqrt(1-exp(-2*a*delta))*rnorm(1,0,1)
}

r_T <-function(a,b,sigma,delta,N){
  r <- 1
  data <- matrix(nrow = N+1)
  data[1] <- 1
  for(i in 1:N){
    r <- data[i+1] <- r_t(r,a,b,sigma,delta)
  }
  data
}

a_data <- c(0.1,1,5)
sigma_data <- c(0.1,0.2,0.5)
b_data <- c(-5,5)
library(ggplot2)

result <- array(0,c(3,3,2,250001))
for(i in 1:3){
  a <- a_data[i]
  for(j in 1:3){
    sigma <- sigma_data[j]
    for(k in 1:2){
      b <- b_data[k]
      result[i,j,k,] <- r_T(a,b,sigma,1/500,250000)
    }
  }
}
```
Thus, the graph will looks like

```{r}
plot1 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,1,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.1,b=-5")
plot1
plot2 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,1,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.1,b=-5")
plot2
plot3 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,1,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.1,b=-5")
plot3
plot4 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,2,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.2,b=-5")
plot4
plot5 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,2,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.2,b=-5")
plot5
plot6 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,1,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.2,b=-5")
plot6
plot7 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,3,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.5,b=-5")
plot7
plot8 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,3,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.5,b=-5")
plot8
plot9 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,3,1,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.5,b=-5")
plot9

plot11 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,1,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.1,b=5")
plot11
plot12 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,1,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.1,b=5")
plot12
plot13 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,1,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.1,b=5")
plot13
plot14 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,2,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.2,b=5")
plot14
plot15 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,2,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.2,b=5")
plot15
plot16 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,1,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.2,b=5")
plot16
plot17 <- ggplot(data.frame(N = seq(1:250001),r_t = result[1,3,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=0.1,sigma=0.5,b=5")
plot17
plot18 <- ggplot(data.frame(N = seq(1:250001),r_t = result[2,3,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=1,sigma=0.5,b=5")
plot18
plot19 <- ggplot(data.frame(N = seq(1:250001),r_t = result[3,3,2,]),aes(x = N, y = r_t))+
        geom_point(col="steelblue", size=0.1)+
        labs(title="Sample Path", subtitle="a=5,sigma=0.5,b=5")
plot19




```

## 3. Use the Euler–Maruyama method (or the Euler method; see Wiki) to approximate a simulation from the process. Specifically, partition the time interval into a grid with subintervals of equal length $\delta > 0$ for a small $\delta$; approximate $r(t + \delta)$ by a normal random variable with mean $r(t) + \alpha(b - r(t)) \delta$ and standard deviation $\sigma \delta$. Write a function to implement this approximation with $\delta$ as one of the arguments. For $\delta \in \{1, 0.5, 0.1, 0.01\}$, generate a sample of size 1000 for $r(1)$. Plot the kernel densities against the true density.

Set $r(0)=1$, $\alpha = 0.1$, $\sigma = 0.1$, $b = -5$

```{r}
r_t1 <- function(r,a,b,sigma,delta){
  rnorm(1,r + a*(b-r)*delta,sigma*delta)
}

r_T1 <-function(a,b,sigma,delta){
  r <- 1
  length <- 1/delta
  for(i in 1:length){
    r <-  r_t1(r,a,b,sigma,delta)
  }
  r
}

a <- 0.1
sigma <- 0.1
b <- -5
delta_data <- c(1,0.5,0.1,0.01)
result <- matrix(nrow = 4,ncol = 1000)
for(i in 1:4){
  delta <- delta_data[i]
  for(j in 1:1000){
    result[i,j] <- r_T1(a,b,sigma,delta)
  }
}

library(ggplot2)
```

```{r plot1, echo=FALSE}
plot1<- ggplot(data.frame(x = result[1,]), aes(x = x))+ 
  geom_density(colour ='steelblue',alpha=0.2)+
  geom_histogram(aes(y=..density..))+labs(x = 'r(t)', 
  y = 'density', title='Kernel Density', subtitle = 'delta = 1')
plot1
plot2<- ggplot(data.frame(x = result[2,]), aes(x = x))+ 
  geom_density(colour ='steelblue',alpha=0.2)+
  geom_histogram(aes(y=..density..))+labs(x = 'r(t)', 
  y = 'density', title='Kernel Density', subtitle = 'delta = 0.5')
plot2
plot3<- ggplot(data.frame(x = result[3,]), aes(x = x))+ 
  geom_density(colour ='steelblue',alpha=0.2)+
  geom_histogram(aes(y=..density..))+labs(x = 'r(t)', 
  y = 'density', title='Kernel Density', subtitle = 'delta = 0.1')
plot3
plot4<- ggplot(data.frame(x = result[4,]), aes(x = x))+ 
  geom_density(colour ='steelblue',alpha=0.2)+
  geom_histogram(aes(y=..density..))+labs(x = 'r(t)', 
  y = 'density', title='Kernel Density', subtitle = 'delta = 0.01')
plot4
```
and The true value of r(1) can be calculated by 
```{r}
r_t2 <- function(r,a,b,sigma,delta){
  exp(-a*delta)*r+b*(1-exp(-a*delta))+sigma/sqrt(2*a)*sqrt(1-exp(-2*a*delta))*rnorm(1,0,1)
}

a <- 0.1
sigma <- 0.1
b <- -5
delta <- 1
result <- matrix(nrow =1,ncol = 1000)
for(i in 1:1000){
  result[1,i] <- r_t2(1,a,b,sigma,delta)
}
plot5 <- ggplot(data.frame(x = result[1,]), aes(x = x))+ 
  geom_density(colour ='steelblue',alpha=0.2)+
  geom_histogram(aes(y=..density..))+labs(x = 'r(t)', 
  y = 'density', title='True Density')
plot5
```


# 5.3.4 Poisson Process

## let $\lambda(t) = \sqrt{t} + e^{-t} \sin(2 \pi t)$ be the inensity function of possion process over $t \in [0, 5]$. Let M(t) be the number of events by time t.

## 1. what is the distribution of N(5) and its parameter(s)? Use Mathematica or Maple for integration if needed.

As we know that the intensity function of Poisson process over $t \in [0, 5]$ is $\lambda(t) = \sqrt{t} + e^{-t} \sin(2 \pi t)$. At time let, let $N(t)$ be the number of events.  

By \[N(t+s)-N(t) \sim Poisson(\int_{t}^{t+s}\lambda(t)dt) \], we can have 
\[N(5) \sim Poisson(\int_{0}^{5}(\sqrt{t} + e^{-t} \sin(2 \pi t) )dt)\]

Thus, to calculate the integration we have:

```{r}
lam <- function (x){
  sqrt(x) + exp(-x) *sin(2*pi*x)
}

integrate(lam,0,5)$value
```
```{r}
lambda1 <- 7.607738
library(ggplot2)
plot <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))+
  labs(x = 'x', y = 'Density', title='Distribution of N(5)', subtitle = 'lambda = 7.6077')
fun.1 <- function(x) dpois(x,lambda = lambda1)
plot + stat_function(fun = fun.1,col="steelblue") + xlim(0,100)
```

## 2. write a function to simulate from this Poisson process.

To simulation of event times of a non homegeneous Poisson process with rate $\lambda(t)$ to time T, first for all $t \leq T$, we need to consider $\lambda$ such that $\lambda(t) \leq \lambda$ and at $t=0$,$k=0$, draw $r ~U(0,1)$. Then let $t=t-ln(r)/\lambda$, if we have $t>T$, stop and generate $s ~U(0,1)$, if we have $s\leq \frac{\lambda(t)}{\lambda}$ then $k=k+1$, $S(k)=t then we need to go back to draw $r ~U(0,1)$ again.

```{r}
simulate <- function(lambda,T){
  t <-0
  k <-0
  sample <- double(0)
  while(t < T) {
    u <- runif(1,0,1)
    if (u <= lam(t)/lambda) { 
      sample <- append(sample, t)
      k <- k + 1
    }
    t <- t - log(runif(1,0,1))/lambda
  }
  return(sample)
}
result <- double(10000)
for(i in 1:10000){
  result[i] <- length(simulate(15,5))
}
mean(result)

```
We choose the $\lambda=15$ , $T=5$ and simulate for 10000 times and the mean value is about 7.6 which is close to the integration result 7.6077.

## 3. Generate events from this Poisson process 1000 times. Pool all the event points together as a sample and plot their kernel density. Overlay $\lambda(t)/\int_{0}^{5}\lambda(s)ds$ with the kernel density.

```{r}
result <- double(0)
for(i in 1:1000){
  result <- append(result,simulate(15,5))
}
plot1 <- ggplot(data.frame(x = result), aes(x = x))+ 
  geom_density(colour ='blue',alpha=0.2)+
  labs(x = 'r(t)',y = 'Density', title='Kernel and true density')
fun.1 <- function(x) (sqrt(x) + exp(-x)*sin(2*pi*x))/7.607738
plot1 + stat_function(fun = fun.1,col="red") + xlim(0,5)
```
 
 In the graph, the blue line is kernel density and red line is the true density.









